BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-09 LMThesisNicchi.tex
Abstract revised

\newcommand{\bluepill}{{\sc BluePill}}
\newcommand{\mynote}[1]{\marginpar{#1}}

@@

As a multitude of new malware instances arise every day, dynamic analysis techniques have assumed a key role in the detection and inspection of malicious software. One of the most adopted dynamic analysis strategies is function call monitoring, since the abstractions embodied by API calls and system calls can be used to comprehend the visible behaviour of the malware sample under analysis, regardless of any obfuscations made on its code. Function call monitoring is typically implemented by means of API hooking, i.e., the interception of function calls made available by dynamically linked libraries (DLL). However, all dynamic analyses techniques - including API hooking - have to face the adversarial behaviour of \textit{evasive malware} strains, which check if they are being executed in a hostile environment and conceal their malicious behaviour accordingly. Furthermore, the majority of API hooking techniques proposed to date in literature are easily detectable by malware and are not equipped with any cloaking method to hide their presence from evasive malware.

We propose a robust library and system call tracer for Windows applications, specialized in evasive malware. To implement our tool, we used the Intel Pin dynamic instrumentation framework, which allows to transparently inject analysis code in a running binary program. In addition, our tool was integrated with BluePill, a stealthy execution framework developed at Sapienza University and Royal Holloway university, enabling it to counteract a wide range of evasive techniques. Our tracer is capable of tracing input parameters, output buffers and return values of a plethora of system calls (including Windows Native APIs) and library calls, as well as the occurrence of user-mode callbacks and Windows Asynchronous Procedure Calls. We first evaluated our tool on a set of benign applications performing a variety of tasks and exercising a large number of functions. Then, we validated it through the use of a popular security product aimed at assessing a how stealthy a malware analysis system is with respect to a large portion of public evasion techniques employed by real malware families.

>>>>

\newcommand{\bluepill}{{\sc BluePill}}
\newcommand{\mynote}[1]{\marginpar{\color{red} #1}}

@@

As a multitude of new malware instances arise every day, dynamic analysis techniques have assumed a key role in the detection and inspection of malicious software. One of the most adopted dynamic analysis strategies is function call monitoring, since the abstractions embodied by API calls and system calls can be used to comprehend the visible behaviour of the malware sample under analysis, regardless of any obfuscations made on its code. Function call monitoring is typically implemented by means of API hooking, i.e., the interception of function calls made available by dynamically linked libraries (DLL). However, all dynamic analyses techniques - including API hooking - have to face the adversarial behaviour of \textit{evasive malware} strains, which check if they are being executed in a hostile environment and conceal their malicious behaviour accordingly. Furthermore, the majority of API hooking techniques proposed to date in literature are detectable, as they are not equipped with cloaking methods to hide their presence from evasive malware.

We propose a robust library and system call tracer for Windows applications, specialized in evasive malware. To implement our tool, we used the Intel Pin dynamic instrumentation framework, which allows transparent injection of analysis code in a running binary program. In addition, our tool was integrated with BluePill, a stealthy execution framework developed at Sapienza University and Royal Holloway university, enabling BlueTracer to counteract a wide range of evasive techniques.

Our tracer is capable of tracing input parameters, output buffers and return values of a plethora of system calls (including Windows Native APIs) and library calls, as well as the occurrence of user-mode callbacks and Windows Asynchronous Procedure Calls. We first evaluated our tool on a set of benign applications performing a variety of tasks and exercising a large number of functions. Then, we validated it through the use of a popular security product aimed at assessing how stealthy a malware analysis system is with respect to a large body of public evasion techniques employed by real malware families seen in the wild.
